---
title: "draft.2020.07"
author: "Zongchao Liu"
date: "6/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
library(tidyverse)
library(caret)
library(ModelMetrics)
library(glmnet)
```


```{r,message=FALSE}
train = read_csv('../data_checking/train101.csv') %>% .[,-1]
test = read_csv('../data_checking/test101.csv') %>% .[,-1]
```

# calculate correlation / avoid overfitting

```{r}
library(mlbench)
library(ranger)
# calculate correlation matrix
correlationMatrix <- cor(scale(train[,3:1220]))
# summarize the correlation matrix
#print(correlationMatrix)
# find attributes that are highly corrected (ideally >0.75)
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff = 0.85)
# print indexes of highly correlated attributes
#print(highlyCorrelated)
1218 - length(highlyCorrelated)
```


```{r}
library(MLmetrics)
#features.lasso = features %>% .[features_into_account] #datase for lasso
#x = features.lasso %>% as.matrix() %>% scale()
#y = features$pcr %>% as.factor()

features.lasso = train %>%
  .[,-c(1:2)] %>%
  .[,-highlyCorrelated] %>%
  mutate(pcr = train$pcr) %>% 
  select(pcr,everything())

x = features.lasso %>%  .[,-1] %>% as.matrix() %>% scale()
y = train$pcr %>% as.factor()

levels(y) = c("N","Y")

ctrl1  = trainControl(method = "cv", number = 10,
                      summaryFunction = prSummary,
                       # prSummary needs calculated class probs
                       classProbs = T)

set.seed(123)
lasso.fit = train(x, y,method = "glmnet",
                  tuneGrid = expand.grid(alpha = 1,lambda = seq(0.045, 0.07, length = 100)),
        #preProc = c("center", "scale"), #0.55，0.08
        #metric = "ROC",
        trControl = ctrl1)
lasso.fit$bestTune
#lasso.fit$finalModel$beta
ggplot(lasso.fit,highlight = T) +
  theme_bw()
  labs(title = "Variable Selection Process via Logistic LASSO Regression") +
  theme(plot.title = element_text(hjust = .5))

#trace plots for lasso variable selection
library(glmnet)

plot.glmnet(lasso.fit$finalModel, xvar = "lambda", label = T,
     main = "Trace Plot for the Coefficients")

lasso.fit$bestTune
```

# obtain the selected features

```{r}
coef = data.frame(as.matrix(coef(lasso.fit$finalModel,lasso.fit$bestTune$lambda)))

coef = coef %>%
  mutate(coef = rownames(coef)) %>%
  rename("value" = "X1") %>%
  filter(value !=0)

coef_num = coef %>%
  filter(value !=0) %>%
  nrow() 

print(coef_num)
```

# calculate radscore

```{r}
# The selected predictors are:
predictors = coef$coef[-1] #plus intercept 1+12
# the next step is to calculate radscores:
predictors_val = coef$value[-1]

# the next step is to calculate radscores:
feature.matrix = train[predictors] %>% as.matrix() %>% scale() 
coef.matrix = predictors_val %>% as.matrix()
radscore = feature.matrix %*% coef.matrix + coef$value[1]

# construct a new dataset for future prediction
data.pred = train %>%
  mutate(radscore = radscore) %>%
  mutate(pcr = factor(pcr))
```


# plot out scores

```{r}
# train
data.pred %>%
  mutate(id = 1:nrow(data.pred)) %>%
  ggplot(aes(x = pcr,
             y = radscore, fill = pcr)) +
  geom_boxplot() +
  theme_bw()


data.pred %>%
  arrange(pcr) %>%
  mutate(id = 1:nrow(data.pred)) %>%
  mutate(radscore = radscore+1) %>%
  ggplot(aes(x = reorder(id,radscore), y = radscore,
              fill = pcr, vjust = ifelse(radscore > 0, 0, 1))) +
  geom_col() +
  geom_hline(yintercept = 0) +
  scale_y_continuous(labels = function(x) x - 1) +
  theme_classic() +
  labs(title = "Radscores for the Primary",
       x = "Patients",
       y = "Radscore") + 
  theme(plot.title = element_text(hjust = .5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggsci::scale_fill_jama(labels = c("PCR = 0", "PCR = 1"))

ggsave("../data_checking/primary_score.pdf")
dev.off()
  
  


# validation
# the next step is to calculate radscores:
feature.matrix.val = test[predictors] %>% as.matrix() %>% scale()
coef.matrix = predictors_val %>% as.matrix()
radscore.val = feature.matrix.val %*% coef.matrix + coef$value[1]

# construct a new dataset for future prediction
data.pred.val = test %>%
  mutate(radscore = radscore.val,
         pcr = factor(pcr))


# plot
data.pred.val %>%
  arrange(pcr) %>%
  mutate(id = 1:nrow(data.pred.val)) %>%
  mutate(radscore = radscore+1) %>%
  ggplot(aes(x = reorder(id,radscore), y = radscore,
              fill = pcr, vjust = ifelse(radscore > 0, 0, 1))) +
  geom_col() +
  geom_hline(yintercept = 0) +
  scale_y_continuous(labels = function(x) x - 1) +
  theme_classic() +
  labs(title = "Radscores for the 58 Patients",
       x = "Patients",
       y = "Radscore") + 
  theme(plot.title = element_text(hjust = .5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggsci::scale_fill_jama(labels = c("PCR = 0", "PCR = 1"))
ggsave("../data_checking/val_score.pdf")
dev.off()

```


# prepare for modeling


```{r,message=FALSE}
scores = rbind(data.pred[,c(1,2,1221)],data.pred.val[,c(1,2,1221)]) %>%
  rename("Name" = "name")

features = read_csv('../2020.05/feature_extracted_total177.csv')
clinical = read_csv('../2020.05/outcome_2020_total(1).csv') %>%
  mutate(Name = features$name)
rm(features)

data_full = left_join(clinical,scores, by = "Name") %>% select(-pcr)
train_full = data_full %>% filter(Name %in% train$name)
test_full = data_full %>% filter(Name %in% train$name == FALSE)

train_full = train_full %>%
  mutate(Age = factor(Age),
         cT = ifelse(cT == 2, 3, cT),
         cT = factor(cT),
         cN = factor(cN),
         cTNM = factor(cTNM),
         MRF = as.character(MRF),
         `Tumor length` = factor(`Tumor length`),
         Distance = factor(Distance),
         CEA = factor(CEA),
         Differentiation = factor(Differentiation),
         Group = ifelse(Group == "C", "B", Group),
         PCR = factor(PCR)
         #rf.score = rf.score
         )

test_full = test_full %>%
  mutate(Age = factor(Age),
         cT = ifelse(cT == 2, 3, cT),
         cT = factor(cT),
         cN = factor(cN),
         cTNM = factor(cTNM),
         MRF = as.character(MRF),
         `Tumor length` = factor(`Tumor length`),
         Distance = factor(Distance),
         CEA = factor(CEA),
         Differentiation = factor(Differentiation),
         Group = ifelse(Group == "C", "B", Group),
         PCR = factor(PCR)
         #rf.score = rf.score
         )
```



# modeling

# reg

```{r,message=FALSE,warning=FALSE}
train_full = train_full %>% mutate(radscore = as.numeric(train_full$radscore))
test_full = test_full %>% mutate(radscore = as.numeric(test_full$radscore))


set.seed(123)
ctrl = trainControl(method = "repeatedcv", number = 10, repeats = 5,
                    classProbs = TRUE,
                    summaryFunction = twoClassSummary,
                    sampling = "up")

levels(train_full$PCR) = c("N", "Y")
levels(test_full$PCR) = c("N", "Y")

set.seed(123)
rad.fit = train(x = train_full[,c(3,6,8,10,12,13,15)], # radscore and others
                y = train_full$PCR,
                method = "glm",
                metric = "ROC",
                trControl = ctrl
                ) #radscore and others (logistic

set.seed(123)
x.lasso = model.matrix(PCR ~. ,train_full[,c(2:3,5:15)])
y.lasso = train_full[,c(2:15)]$PCR

#lasso.predict.fit = train(x = x.lasso, # radscore with others (lasso)
#                y = y.lasso,
#                method = "glmnet",
#                tuneGrid = expand.grid(alpha = 1,lambda = exp(seq(-5, 5, length=1000))),
#                preProc = c("center", "scale"),
#                trControl = ctrl)

#rad_and_group.fit = train(x = train_full[,c(4,15)], # only radscore and group(logistic)
#                y = train_full$PCR,
#                method = "glm",
#                metric = "ROC",
#                trControl = ctrl
#                )
set.seed(123)
all_other.fit = train(x = train_full[,c(2:3,5:13)], # only others
                y = train_full$PCR,
                method = "glm",
                metric = "ROC",
                trControl = ctrl
                )

test.pred.prob = predict(rad.fit, newdata = test_full , type = "prob")[,2]
test.pred = rep("N", length(test.pred.prob ))
test.pred[test.pred.prob > 0.5] = "Y"

caret::confusionMatrix(data = factor(test.pred,levels = c("N","Y")),
                       reference = test_full$PCR,
                       positive = "Y")
```



```{r}
#train_full = train_full %>% as.matrix() %>% as.data.frame()
#test_full = test_full %>% as.matrix() %>% as.data.frame()


#rf.grid = expand.grid(mtry = 2:9,
#                       splitrule = "gini",
#                       min.node.size = 10:15
#                      )
#set.seed(123)
#rf.fit = train(PCR ~ ., train_full[,-c(1,4)], 
#                #subset = rowTrain,
#                method = "ranger",
#                tuneGrid = rf.grid,
#                metric = "ROC",
#                trControl = ctrl)
#
##rf.fit = train(PCR ~ ., data.pred[rowTrain,-1], 
##                #subset = rowTrain,
##                method = "rf",
##                ntree = 1000,
##                tuneGrid = data.frame(mtry = 1:6),
##                metric = "ROC",
##                trControl = ctrl)
#
#
#
#ggplot(rf.fit, highlight = TRUE) +
#  theme_minimal() +
#  labs(title = "Tuning Parameters for Random Forest") +
#  theme(plot.title = element_text(hjust = .5)) +
#  ggsci::scale_color_lancet()
#rf.fit$bestTune
#
#set.seed(123)
#data.pred.rf = train_full
#colnames(data.pred.rf) = make.names(colnames(data.pred.rf))
#rf.final = ranger(PCR ~ ., 
#                  data.pred.rf[,-c(1,4)], 
#                  mtry = rf.fit$bestTune$mtry, 
#                  splitrule = "gini",
#                  min.node.size = rf.fit$bestTune$min.node.size,
#                  importance = "impurity")
#
## var importance
#vip::vip(rf.final) + theme_bw() + ggtitle("Variable Importance - RF")
#barplot(sort(ranger::importance(rf.final), decreasing = FALSE),
#        las = 2, 
#        horiz = TRUE, 
#        cex.names = 0.6,
#        col = colorRampPalette(colors = c("black","gold"))(20))
#
#
## explain prediction
## fixed a small bug
library(lime)
new_obs = train_full[,-c(1,4)][c(1,28),]
#explainer.rf = lime(train_full[,-c(1,4)], rf.fit)
#explanation.rf = lime::explain(new_obs, explainer.rf, n_features = 8, labels = "Y")
#
#plot_features(explanation.rf)
#plot_explanations(explanation.rf)
#attention!!! => actual result: case 1 is Y, case 2 is N
```
# SVm

```{r}
sum_y = sum(train_full$PCR == "Y")
ratio = sum(train_full$PCR == "Y")/length(train_full$PCR)
weights_class = ifelse(train_full$PCR == "Y", 0.1,0.9 )

set.seed(123)
svm.fit = train(PCR ~ ., train_full[,-c(1,4)] ,
                method = "svmLinear2", 
                trControl = ctrl,  
                #preProcess = c("center","scale"),
                metric = "ROC",
                #weights = weights_class,
                tuneGrid = expand.grid(cost = seq(1,5, length = 20))
                )


ggplot(svm.fit,highlight = TRUE) + theme_bw()
```

# boosting

```{r}
gbm.grid = expand.grid(n.trees = c(750,1250,1500),
                        interaction.depth = 1,
                        shrinkage = c(0.05),
                        n.minobsinnode = 3)
set.seed(123)
# Binomial loss function
gbm.fit = train(PCR ~ ., train_full[,-c(1,4)],
                 tuneGrid = gbm.grid,
                 trControl = ctrl,
                 method = "gbm",
                 distribution = "bernoulli",
                 metric = "ROC",
                 verbose = FALSE)

ggplot(gbm.fit, highlight = T) +
  theme_minimal() +
  labs(title = "Tuning Parameters for Gradient Boosting Model") +
  theme(plot.title = element_text(hjust = .5)) #+
  #ggsci::scale_color_lancet()
gbm.fit$bestTune

vip::vip(gbm.fit$finalModel) + theme_bw() + ggtitle("Variable Importance - #GBM")
summary(gbm.fit$finalModel, las = 2, cBars = 19, cex.names = 0.6)

explainer.gbm = lime(train_full[,-c(1,4)], gbm.fit)
explanation.gbm = explain(new_obs, explainer.gbm, n_features = 8, labels = "Y")
plot_features(explanation.gbm)

```


# resample

```{r}
res = resamples(list(logistic_all = rad.fit,
                     #rad_and_group = rad_and_group.fit,
                     all_other = all_other.fit,
                     #rf = rf.fit,
                     #lasso = lasso.predict.fit,
                     GBM = gbm.fit,
                     svm = svm.fit),
                     metric = "ROC")
summary(res)
bwplot(res)
```



# table 1 grouped by radscores

```{r,eval=FALSE}
#library(kableExtra)
#label = rep("Primary Cohort",177)
#label[-rowTrain] = "Validation Cohort"
#data.tbl = data.tbl %>%
#  mutate(cohort = label)
#
#primary = data.tbl[rowTrain,]
#validation = data.tbl[-rowTrain,]
#
#
#control = arsenal::tableby.control(test = T,
#                                    numeric.stats = c("meansd","medianq1q3","range"),
#                                    #cat.stats = c("countrowpct"),
#                                    ordered.stats = c("Nmiss"),
#                                    digits = 2)
#
#tbl_primary = arsenal::tableby(PCR~ Age + Sex + Group + cT + cN +cTNM + MRF + `Tumor length` + #`Tumor thickness` + Distance + CEA  + Differentiation + radscore, data = primary,control = #control)
#
#tbl_validation = arsenal::tableby(PCR~ Age + Sex + Group + cT + cN +cTNM + MRF + `Tumor length` #+ `Tumor thickness` + Distance + CEA  + Differentiation + radscore, data = validation, control #= control)
#
#summary(tbl_primary,text = T) %>%
#  knitr::kable(booktabs = T, caption = "Descriptive Statistics") %>%
#  kable_styling(latex_options = c("striped", "hold_position"))
#  
#  
#summary(tbl_validation,text = T) %>%
#  knitr::kable(booktabs = T, caption = "Descriptive Statistics") %>%
#  kable_styling(latex_options = c("striped", "hold_position"))
```


# Making ROC curves for logistic regression Models

## for training set

```{r}
library(pROC)
# rad.fit ROC (radscore and others)
rad.fit.pred.train.prob = predict(rad.fit,newdata = train_full,type = "prob")[,2]# select pos resp
rad.fit.pred.train = rep("N",length(rad.fit.pred.train.prob))
rad.fit.pred.train[rad.fit.pred.train.prob > 0.5] = "Y"

roc.rad.fit = roc(train_full$PCR, rad.fit.pred.train.prob)
plot(roc.rad.fit,
     legacy.axes = T,
     print.auc = T,
     print.thres = T)
ci.auc(roc.rad.fit)
#plot(smooth(rad_and_group.fit.fit),
     #add = T,
     #col = 4)
## only radscore and groups
#rad_group.fit.pred.train.prob = predict(rad_and_group.fit ,newdata = train_full,type = #"prob")[,2]# select pos resp
#rad_group.fit.pred.train = rep("N",length(rad_group.fit.pred.train.prob))
#rad_group.fit.pred.train[rad_group.fit.pred.train.prob > 0.5] = "Y"

#roc.rad_group.fit = roc(train_full$PCR, rad_group.fit.pred.train.prob)
#plot(roc.rad_group.fit,
#     legacy.axes = T,
#     print.auc = T,
#     print.thres = T)
#ci.auc(roc.rad_group.fit)

# RF train ROC
#rf.fit.pred.train.prob = predict(rf.fit,newdata = train_full[,-1],type = "prob")[,2]# select #pos resp
#rf.fit.pred.train = rep("N",length(rf.fit.pred.train.prob))
#rf.fit.pred.train[rf.fit.pred.train.prob > 0.5] = "Y"
#
#roc.rf.fit = roc(train_full$PCR, rf.fit.pred.train.prob)
#plot(roc.rf.fit,
#     legacy.axes = T,
#     print.auc = T,
#     print.thres = T)
#ci.auc(roc.rf.fit)

# Boosting train ROC
gbm.fit.pred.train.prob = predict(gbm.fit, newdata = train_full[,-1], type = "prob")[,2]# select pos resp
gbm.fit.pred.train = rep("N",length(gbm.fit.pred.train.prob))
gbm.fit.pred.train[gbm.fit.pred.train.prob > 0.5] = "Y"

roc.gbm.fit = roc(train_full$PCR, gbm.fit.pred.train.prob)
plot(roc.gbm.fit,
     legacy.axes = T,
     print.auc = T)
ci.auc(roc.gbm.fit)

# SVM
svm.fit.pred.train.prob = predict(svm.fit, newdata = train_full[,-1], type = "prob")[,2]# select pos resp
svm.fit.pred.train = rep("N",length(svm.fit.pred.train.prob))
svm.fit.pred.train[svm.fit.pred.train.prob > 0.5] = "Y"

roc.svm.fit = roc(train_full$PCR, svm.fit.pred.train.prob)
plot(roc.svm.fit,
     legacy.axes = T,
     print.auc = T)
ci.auc(roc.svm.fit)


plot(roc.rad.fit, print.auc=TRUE,print.auc.x=0.3,print.auc.y=0.1,auc.polygon.col="gray", grid = c(0.5, 0.3), max.auc.polygon=TRUE,legen = T,main = "Primary Cohort")
#plot.roc(roc.rf.fit,add=T,col="red", print.auc=TRUE,print.auc.x=0.3,print.auc.y=0.2)
plot.roc(roc.svm.fit,add=T,col="blue",print.auc=TRUE,print.auc.x=0.3,print.auc.y=0.3,)
plot.roc(roc.gbm.fit,add=T,col="yellow",print.auc=TRUE,print.auc.x=0.3,print.auc.y=0.4)
legend("bottomright",legend = c("radscore with other clinical variables",
                  "random forest",
                  "gradient boosting model",
                  "SVM"),
       col = c("black","red","yellow","blue"),lwd = 4,cex = 0.35)
```

## for validation set 

```{r}
# rad.fit ROC (radscore and others)
rad.fit.pred.test.prob = predict(rad.fit,newdata = test_full,type = "prob")[,2]# select pos resp
rad.fit.pred.test = rep("N",length(rad.fit.pred.test.prob))
rad.fit.pred.test[rad.fit.pred.test.prob > 0.5] = "Y"

roc.rad.fit.test = roc(test_full$PCR, rad.fit.pred.test.prob)
plot(roc.rad.fit.test,
     legacy.axes = T,
     print.auc = T,
     print.thres = T)
ci.auc(roc.rad.fit.test)
#plot(smooth(rad_and_group.fit.fit),
#     add = T,
#     col = 4)

# only radscore and groups
#rad_group.fit.pred.test.prob = predict(rad_and_group.fit ,newdata = test_full,type = #"prob")[,2]# select pos resp
#rad_group.fit.pred.test = rep("N",length(rad_group.fit.pred.test.prob))
#rad_group.fit.pred.test[rad_group.fit.pred.test.prob > 0.5] = "Y"
#
#roc.rad_group.fit.test = roc(test_full$PCR, rad_group.fit.pred.test.prob)
#plot(roc.rad_group.fit.test,
#     legacy.axes = T,
#     print.auc = T,
#     print.thres = T)
#ci.auc(roc.rad_group.fit.test)

## RF train ROC
#rf.fit.pred.test.prob = predict(rf.fit,newdata = test_full[,-1],type = "prob")[,2]# select #pos resp
#rf.fit.pred.test = rep("N",length(rf.fit.pred.test.prob))
#rf.fit.pred.test[rf.fit.pred.test.prob > 0.5] = "Y"
#
#roc.rf.fit.test = roc(test_full$PCR, rf.fit.pred.test.prob)
#plot(roc.rf.fit.test,
#     legacy.axes = T,
#     print.auc = T,
#     print.thres = T)
#ci.auc(roc.rf.fit.test)

# Boosting test ROC
gbm.fit.pred.test.prob = predict(gbm.fit,newdata = test_full[,-1], type = "prob")[,2]# elect #pos resp
gbm.fit.pred.test = rep("N",length(gbm.fit.pred.test.prob))
gbm.fit.pred.test[gbm.fit.pred.test.prob > 0.5] = "Y"

roc.gbm.fit.test = roc(test_full$PCR, gbm.fit.pred.test.prob)
plot(roc.gbm.fit.test,
     legacy.axes = T,
     print.auc = T)
ci.auc(roc.gbm.fit.test)

# Boosting test ROC
svm.fit.pred.test.prob = predict(svm.fit,newdata = test_full[,-1], type = "prob")[,2]# elect #pos resp
svm.fit.pred.test = rep("N",length(svm.fit.pred.test.prob))
svm.fit.pred.test[svm.fit.pred.test.prob > 0.5] = "Y"

roc.svm.fit.test = roc(test_full$PCR, svm.fit.pred.test.prob)
plot(roc.svm.fit.test,
     legacy.axes = T,
     print.auc = T)
ci.auc(roc.svm.fit.test)




plot(roc.rad.fit.test, print.auc=TRUE,print.auc.x = 0.3,print.auc.y= 0.1,grid = c(0.5, 0.3), max.auc.polygon=TRUE,legen = T,main = "Validation Cohort")
#plot.roc(roc.rf.fit.test,add=T,col="red", print.auc=TRUE,print.auc.x=0.3,print.auc.y=0.2)
plot.roc(roc.svm.fit.test,add=T,col="blue",print.auc=TRUE,print.auc.x=0.3,print.auc.y=0.3,)
plot.roc(roc.gbm.fit.test,add=T,col="yellow",print.auc=TRUE,print.auc.x=0.3,print.auc.y=0.4)
legend("bottomright",legend = c("radscore with other clinical variables-LR",
                  "SVM",
                  "Randomforest",
                  "GBM"
                  ),
       col = c("black","blue","red","yellow"),lwd = 4,cex = 0.35)
```


# logsitic regression's caliberation & HL test

```{r}
##?calibration()
#trainProbs = data.frame(
#obs = train_full$PCR,
#rad = 1-rad.fit.pred.train.prob,
#rf = 1-rf.fit.pred.train.prob,
#rad_and_group = 1-rad_group.fit.pred.train.prob,
#gbm = 1-gbmB.fit.pred.train.prob)
#
#testProbs = data.frame(
#  obs = train_full$PCR,
#  rad = 1-rad.fit.pred.test.prob,
#  rf = 1-rf.fit.pred.test.prob,
#  rad_and_group = 1-rad_group.fit.pred.test.prob,
#  gbm = 1-gbmB.fit.pred.test.prob
#)
#
#cal.int = calibration(obs ~ rad +
#                        rf + 
#                        rad_and_group + gbm,
#                        data = trainProbs)
#cal.ext = calibration(obs ~ rad + 
#                        rf + rad_and_group +gbm 
#                      , data = testProbs)
#
#plot(cal.int,type = "l",auto.key = list(columns = 4,
#                                          lines = TRUE,
#                                          points = FALSE),
#     main = "Calibration Curves for Primary Cohort")
#
#plot(cal.ext,type = "l",auto.key = list(columns = 4,
#                                          lines = TRUE,
#                                          points = FALSE),
#     main = "Calibration Curves for Validation Cohort")


```

# results of the logsitic regression
```{r}
### radscore and other variables
##summary(rad.fit$finalModel)
### radscore and group
##summary(rad_and_group.fit$finalModel)


```

# Validation Performance

```{r}
library(epiR)
# validation results
caret::confusionMatrix(factor(rad.fit.pred.test),reference = test_full$PCR,
                       positive = "Y")
#caret::confusionMatrix(factor(rf.fit.pred.test),reference = test_full$PCR,
#                       positive = "Y")
caret::confusionMatrix(factor(svm.fit.pred.test),reference = test_full$PCR,
                      positive = "Y")
caret::confusionMatrix(factor(gbm.fit.pred.test),reference = test_full$PCR,
                       positive = "Y")
##logistic
#dat1 = as.table(matrix(c(36,1,14,8),nrow = 2, byrow = TRUE))
#rval = epi.tests(dat1, conf.level = .95)
#print(rval)
#summary(rval)
##fr
#caret::confusionMatrix(factor(rf.fit.pred.test),reference = test_full$PCR,
#                       positive = "Y")
#dat2 = as.table(matrix(c(43,2,7,7),nrow = 2, byrow = TRUE))
#rval = epi.tests(dat2,conf.level = .95)
#print(rval)
#summary(rval)
#
##svm
#dat3 = as.table(matrix(c(38,1,12,8),nrow = 2, byrow = TRUE))
#rval = epi.tests(dat3,conf.level = .95)
#print(rval)
#summary(rval)
#
#dat4 = as.table(matrix(c(38,2,12,7),nrow = 2, byrow = TRUE))
#rval = epi.tests(dat4,conf.level = .95)
#print(rval)
#summary(rval)


#http://vassarstats.net/clin1.html#return
```

# Training Performance

```{r}
# primary results
caret::confusionMatrix(factor(rad.fit.pred.train),reference = train_full$PCR,
                       positive = "Y")
#caret::confusionMatrix(factor(rad_group.fit.pred.train),reference = train_full$PCR,
#                       positive = "Y")
caret::confusionMatrix(factor(rf.fit.pred.train),reference = train_full$PCR,
                       positive = "Y")
caret::confusionMatrix(factor(svm.fit.pred.train),reference = train_full$PCR,
                       positive = "Y")
```


# heatmap

```{r,message=FALSE}
pcr_total = rbind(train,test) %>% .[,c(1,2)]
feature_total = rbind(train,test) %>%
  .[coef$coef[-1]] %>%
  scale(.) %>%
  as.data.frame()
feature_total = cbind(pcr_total,feature_total)


# geom_tile:

feature_total %>%
  arrange(pcr) %>%
  mutate(id = factor(1:177)) %>%
  pivot_longer(., cols = 3:12, names_to = "feature") %>%
  ggplot(aes(x = id, y = feature, fill = value)) +
  geom_tile(aes(fill = value,
                width = .9,
                height =.9),
            size = 2.5) + 
  scale_fill_gradientn(colors = c("white","orange","red")) +  
  scale_x_discrete(breaks = c(1:177),
                   labels = as.character(c(1:177)) # 177 patients
                   )

# hestmap
feature_total = feature_total %>%
  arrange(pcr) %>%
  mutate(id = 1:177) %>%
  .[,-2] %>%
  as.matrix()

library(gplots)
library(RColorBrewer)

col1 = colorRampPalette(brewer.pal(9, "GnBu"))(2000)
col2 = colorRampPalette(brewer.pal(8, "Spectral"))(2)
scaleRYG = colorRampPalette(c("red","yellow","darkgreen"), 
                             space = "rgb")(30)

heatmap.2(t(feature_total[,-c(1,ncol(feature_total))]),col = col1,
          keysize = .8,
          key.par = list(cex =.5),
          trace = "none", key = TRUE, cexCol = .75,
          labCol = as.character(feature_total[,ncol(feature_total)]),
          margins =c(10,10),
          Colv = F,
          Rowv = F,
          xlab = "Patient No.",
          ylab = "Features",
          ColSideColors = col2[as.numeric(feature_total[,1])+1],
          #lwid = c(1,1),
          #lhei = c(1,1),
          notecex = 0.1)
  
  


```